Data Exploration

### 2019-01-24 src/data_exploration.Rmd
countpercell
more and more vif monthly

>>>>

As per SLR suggestion, rerun without gear

```{r vif all but gear}
vif_allbutgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutgear, "../output/bio/vif_allbutgear.csv", row.names = TRUE)
vif_allbutgear
```

As per SLR suggestion, rerun without gear and most highly correlated variables

```{r vif without gear + high corr}
vif_allbutgearhighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutgearhighlycorr, "../output/bio/vif_allbutgearhighlycorr.csv", row.names = TRUE)
vif_allbutgearhighlycorr
```

As per SLR suggestion, rerun with gear but without most highly correlated variables

```{r vif all but high corr}
vif_allbuthighlycorr <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter + gear, data = presab))
write.csv(vif_allbuthighlycorr, "../output/bio/vif_allbuthighlycorr.csv", row.names = TRUE)
vif_allbuthighlycorr

vif_allbuthighlycorr_sq <- read.csv("../output/bio/vif_allbuthighlycorr.csv", header = TRUE)
vif_allbuthighlycorr_sq$GVIF2Dfsq <- vif_allbuthighlycorr_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuthighlycorr_sq, "../output/bio/vif_allbuthighlycorr_sq.csv", row.names = TRUE)
vif_allbuthighlycorr_sq
```

ok remove one variable at a time - leave gear in

remove amo_prev
```{r vif all but amoprev}
vif_allbutamo_prev <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutamo_prev, "../output/bio/vif_allbutamo_prev.csv", row.names = TRUE)
vif_allbutamo_prev

vif_allbutamo_prev_sq <- read.csv("../output/bio/vif_allbutamo_prev.csv", header = TRUE)
vif_allbutamo_prev_sq$GVIF2Dfsq <- vif_allbutamo_prev_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutamo_prev_sq, "../output/bio/vif_allbutamo_prev_sq.csv", row.names = TRUE)
vif_allbutamo_prev_sq
```
and leave gear out

remove amo_prev + gear
```{r vif all but amoprev + gear}
vif_allbutamo_prevgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbutamo_prevgear, "../output/bio/vif_allbutamo_prevgear.csv", row.names = TRUE)
vif_allbutamo_prevgear
```

remove chl_depth
```{r vif all but chldepth}
vif_allbutchl_depth <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutchl_depth, "../output/bio/vif_allbutchl_depth.csv", row.names = TRUE)
vif_allbutchl_depth

vif_allbutchl_depth_sq <- read.csv("../output/bio/vif_allbutchl_depth.csv", header = TRUE)
vif_allbutchl_depth_sq$GVIF2Dfsq <- vif_allbutchl_depth_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutchl_depth_sq, "../output/bio/vif_allbutchl_depth_sq.csv", row.names = TRUE)
vif_allbutchl_depth_sq
```

remove chl_depth and gear
```{r vif all but chldepth+gear}
vif_allbutchl_depthgear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutchl_depthgear, "../output/bio/vif_allbutchl_depthgear.csv", row.names = TRUE)
vif_allbutchl_depthgear
```

remove ssh_surface
```{r vif all but ssh}
vif_allbutssh_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutssh_surface, "../output/bio/vif_allbutssh_surface.csv", row.names = TRUE)
vif_allbutssh_surface

vif_allbutssh_surface_sq <- read.csv("../output/bio/vif_allbutssh_surface.csv", header = TRUE)
vif_allbutssh_surface_sq$GVIF2Dfsq <- vif_allbutssh_surface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutssh_surface_sq, "../output/bio/vif_allbutssh_surface_sq.csv", row.names = TRUE)
vif_allbutssh_surface_sq
```

remove ssh_surface & gear
```{r vif all but ssh+gear}
vif_allbutssh_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutssh_surfacegear, "../output/bio/vif_allbutssh_surfacegear.csv", row.names = TRUE)
vif_allbutssh_surfacegear
```

remove o2_surface
```{r vif all but o2surface}
vif_allbuto2_surface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuto2_surface, "../output/bio/vif_allbuto2_surface.csv", row.names = TRUE)
vif_allbuto2_surface

vif_allbuto2_surface_sq <- read.csv("../output/bio/vif_allbuto2_surface.csv", header = TRUE)
vif_allbuto2_surface_sq$GVIF2Dfsq <- vif_allbuto2_surface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuto2_surface_sq, "../output/bio/vif_allbuto2_surface_sq.csv", row.names = TRUE)
vif_allbuto2_surface_sq
```

remove o2_surface & gear
```{r vif all but o2surface+gear}
vif_allbuto2_surfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuto2_surfacegear, "../output/bio/vif_allbuto2_surfacegear.csv", row.names = TRUE)
vif_allbuto2_surfacegear
```

as per SLR chat jan 23:

remove salinity_surface only
```{r vif all but salsurface}
vif_allbutsalinitysurface <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutsalinitysurface, "../output/bio/vif_allbutsalinitysurface.csv", row.names = TRUE)
vif_allbutsalinitysurface

vif_allbutsalinitysurface_sq <- read.csv("../output/bio/vif_allbutsalinitysurface.csv", header = TRUE)
vif_allbutsalinitysurface_sq$GVIF2Dfsq <- vif_allbutsalinitysurface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutsalinitysurface_sq, "../output/bio/vif_allbutsalinitysurface_sq.csv", row.names = TRUE)
vif_allbutsalinitysurface_sq
```
and without gear

```{r vif all but salsurface+gear}
vif_allbutsalinitysurfacegear <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbutsalinitysurfacegear, "../output/bio/vif_allbutsalinitysurfacegear.csv", row.names = TRUE)
vif_allbutsalinitysurfacegear
```

now remove temp_surface plus highly correlated
```{r vif allbut high corr + temp surface/gear/nogear}
#with gear
vif_allbuthighcorrtempsurface <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsurface, "../output/bio/vif_allbuthighcorrtempsurface.csv", row.names = TRUE)
vif_allbuthighcorrtempsurface

vif_allbuthighcorrtempsurface_sq <- read.csv("../output/bio/vif_allbuthighcorrtempsurface.csv", header = TRUE)
vif_allbuthighcorrtempsurface_sq$GVIF2Dfsq <- vif_allbuthighcorrtempsurface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuthighcorrtempsurface_sq, "../output/bio/vif_allbuthighcorrtempsurface_sq.csv", row.names = TRUE)
vif_allbuthighcorrtempsurface_sq

#without gear
vif_allbuthighcorrtempsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_surface + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsurfacegear, "../output/bio/vif_allbuthighcorrtempsurfacegear.csv", row.names = TRUE)
vif_allbuthighcorrtempsurfacegear
```

now remove temp_surface plus salinity_surface
```{r vif allbut temp + salinity surface/gear/nogear}
#with gear
vif_allbuttempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttempsalsurface, "../output/bio/vif_allbuttempsalsurface.csv", row.names = TRUE)
vif_allbuttempsalsurface

vif_allbuttempsalsurface_sq <- read.csv("../output/bio/vif_allbuttempsalsurface.csv", header = TRUE)
vif_allbuttempsalsurface_sq$GVIF2Dfsq <- vif_allbuttempsalsurface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuttempsalsurface_sq, "../output/bio/vif_allbuttempsalsurface_sq.csv", row.names = TRUE)
vif_allbuttempsalsurface_sq

#without gear
vif_allbuttempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_prev + amo_winter, data = presab))
write.csv(vif_allbuttempsalsurfacegear, "../output/bio/vif_allbuttempsalsurfacegear.csv", row.names = TRUE)
vif_allbuttempsalsurfacegear
```

now remove temp_surface plus salinity_surface + highly correlated
```{r vif allbut high corr +  temp + salinity surface/gear/nogear}
#with gear
vif_allbuthighcorrtempsalsurface <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + gear + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsalsurface, "../output/bio/vif_allbuthighcorrtempsalsurface.csv", row.names = TRUE)
vif_allbuthighcorrtempsalsurface

vif_allbuthighcorrtempsalsurface_sq <- read.csv("../output/bio/vif_allbuthighcorrtempsalsurface.csv", header = TRUE)
vif_allbuthighcorrtempsalsurface_sq$GVIF2Dfsq <- vif_allbuthighcorrtempsalsurface_sq$GVIF..1..2.Df..^2
write.csv(vif_allbuthighcorrtempsalsurface_sq, "../output/bio/vif_allbuthighcorrtempsalsurface_sq.csv", row.names = TRUE)
vif_allbuthighcorrtempsalsurface_sq

#without gear
vif_allbuthighcorrtempsalsurfacegear <- vif(lm(occurrence ~ temp_depth + salinity_depth + o2_depth + chl_surface + bottom_depth + mlp_surface + nao_sample + nao_prev + nao_winter + amo_sample + amo_winter, data = presab))
write.csv(vif_allbuthighcorrtempsalsurfacegear, "../output/bio/vif_allbuthighcorrtempsalsurfacegear.csv", row.names = TRUE)
vif_allbuthighcorrtempsalsurfacegear
```

Just out of curiosity, a vif will all but but atmos drivers

```{r}
vif_allbutnaoamo <- vif(lm(occurrence ~ temp_surface + temp_depth + salinity_surface + salinity_depth + o2_surface + o2_depth + chl_surface + chl_depth + bottom_depth + mlp_surface + ssh_surface + gear, data = presab))
write.csv(vif_allbutnaoamo, "../output/bio/vif_allbutnaoamo.csv", row.names = TRUE)
vif_allbutnaoamo

vif_allbutnaoamo_sq <- read.csv("../output/bio/vif_allbutnaoamo.csv", header = TRUE)
vif_allbutnaoamo_sq$GVIF2Dfsq <- vif_allbutnaoamo_sq$GVIF..1..2.Df..^2
write.csv(vif_allbutnaoamo_sq, "../output/bio/vif_allbutnaoamo_sq.csv", row.names = TRUE)
vif_allbutnaoamo_sq
```

# check to see how many cells have >1 point per timeslice

this needs to be 3D so you need to create a cell_id_3d
```{r}
presab$cell_id_3d <- paste(presab$cell_id, presab$depthlayerno, sep="_") #create a new column that is a unique cell_id & depth ID. 
head(presab)
```
and now do one with a timeslice label
```{r}
presab$cell_id_xyzt <- paste(presab$cell_id_3d, presab$year, presab$month, sep="_") #create a new column that is a unique cell_id & depth ID. 
head(presab)
```

the data is in presab$total_cell_obs_xyt 
```{r}
unique_pointpercell <- unique(presab$total_cell_obs_xyzt)
unique_pointpercell
```
bugger... ok so lets do a frequency table
```{r}
celobsmore1 <- subset(presab, total_cell_obs_xyzt > 1)
pointpercell <- with(celobsmore1, table(cell_id_xyzt, total_cell_obs_xyzt))
write.csv(pointpercell, file = "../output/bio/pointpercell.csv")
nrow(pointpercell)
pointpercell
```

so there are 949 3d timesliced cells that contain more than one occurrence. 
- different sampling days
- different coordinates
- risk of duplicate entries (but only if two instituions inputted the same data with slightly different coordinates and different reserch program names - see cell3985_18_2000_7 below as example)

```{r}
cell3985_18_2000_7  <- subset(celobsmore1, cell_id_xyzt == "3985_18_2000_7")
cell3985_18_2000_7 
```

# monthly spearmans correlations and VIF

curious - does correlations/vif alter between months?

First split the dataset into monthly

```{r subset monthly presab}
janpresab <- subset(presab, month == "1")
febpresab <- subset(presab, month == "2")
marpresab <- subset(presab, month == "3")
aprpresab <- subset(presab, month == "4")
maypresab <- subset(presab, month == "5")
junpresab <- subset(presab, month == "6")
julpresab <- subset(presab, month == "7")
augpresab <- subset(presab, month == "8")
seppresab <- subset(presab, month == "9")
octpresab <- subset(presab, month == "10")
novpresab <- subset(presab, month == "11")
decpresab <- subset(presab, month == "12")
```

now get the background points for each month (for spearmans)
```{r monthly background points}
janback <- subset(janpresab, occurrence == "0")
febback <- subset(febpresab, occurrence == "0")
marback <- subset(marpresab, occurrence == "0")
aprback <- subset(aprpresab, occurrence == "0")
mayback <- subset(maypresab, occurrence == "0")
junback <- subset(junpresab, occurrence == "0")
julback <- subset(julpresab, occurrence == "0")
augback <- subset(augpresab, occurrence == "0")
sepback <- subset(seppresab, occurrence == "0")
octback <- subset(octpresab, occurrence == "0")
novback <- subset(novpresab, occurrence == "0")
decback <- subset(decpresab, occurrence == "0")
```

run the correlations on each month's background points

```{r monthly spearman correlations}
janback <- subset(janback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
janback_cor <- cor(janback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(janback_cor, "../output/env/janback_cor.csv", row.names = TRUE)

febback <- subset(febback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
febback_cor <- cor(febback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(febback_cor, "../output/env/febback_cor.csv", row.names = TRUE)

marback <- subset(marback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
marback_cor <- cor(marback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(marback_cor, "../output/env/marback_cor.csv", row.names = TRUE)

aprback <- subset(aprback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
aprback_cor <- cor(aprback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(aprback_cor, "../output/env/aprback_cor.csv", row.names = TRUE)

mayback <- subset(mayback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
mayback_cor <- cor(mayback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(mayback_cor, "../output/env/mayback_cor.csv", row.names = TRUE)

junback <- subset(junback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
junback_cor <- cor(junback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(junback_cor, "../output/env/junback_cor.csv", row.names = TRUE)

julback <- subset(julback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
julback_cor <- cor(julback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(julback_cor, "../output/env/julback_cor.csv", row.names = TRUE)

augback <- subset(augback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
augback_cor <- cor(augback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(augback_cor, "../output/env/augback_cor.csv", row.names = TRUE)

sepback <- subset(sepback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
sepback_cor <- cor(sepback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(sepback_cor, "../output/env/sepback_cor.csv", row.names = TRUE)

octback <- subset(octback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
octback_cor <- cor(octback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(octback_cor, "../output/env/octback_cor.csv", row.names = TRUE)

novback <- subset(novback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
novback_cor <- cor(novback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(novback_cor, "../output/env/novback_cor.csv", row.names = TRUE)

decback <- subset(decback, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
decback_cor <- cor(decback, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(decback_cor, "../output/env/decback_cor.csv", row.names = TRUE)
```

#maxent?

```{r}
library("raster")
library("dismo")
library("rgeos")
```
